---
title: "Market Basket Analysis - Instacart"
output: md_document

---     

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Loading packages
```{r}
suppressMessages(library(dplyr)) 
suppressMessages(library(arules)) 
suppressMessages(library(arulesViz)) 
suppressMessages(library(stringr)) 
```

##Reading and transforming data
First we will read the data and take a brief look at it.
```{r}
setwd("C:/Users/Jimmy Chen/Desktop/Skyrec/FYI/Market Basket")
orders<-read.csv("Market Basket-Instacart.csv")
head(orders)
```

Next we will read the data as a 'transaction' object in order for arules to read it, as well as setting the item info as product name.
```{r}
trans<-read.transactions("Market Basket-Instacart.csv", format = "single", sep=",",cols = c("order_id","product_id"))
orders<-orders[order(as.character(orders$product_id)),]
items<-as.character(unique(orders$product_name)) #get the corresponding name for product id
trans@itemInfo<-data.frame(labels=items)
inspect(trans[1:2])
```

##Item Frequency
Before extracting rules from transaction data, we want to first look at the frequency of which item appeared to gain knowledge of data.
```{r}
itemFrequencyPlot(trans,topN=20,type="absolute")
item_frequency<-itemFrequency(trans,type='absolute')
summary(item_frequency)
```

Some quick thoughts from above:
 - Many of the most popular items are fruits.
 
 - The frequency differs a lot between items, some items are much more popular than others.
 
 - 75% of the items appear equal or less than 5 times, which is a very small ratio of 14000 transactions. in total, this must be considered when deciding minimum support.  
 
##Rules forming
An association rule, for example {diaper->beer} indicates that if a person buys diaper, it is likely to occur that he will also buy beer.The three most frequently used evaluation metric in association rules{X->Y} are:

 - Support : Fraction of transactions that both X and Y appears, using the above example, it means the fraction of transactions that include both diaper and beer.
 
 - Confidence : Given that X is purchased, the conditional probability that Y will also be purchased, for example that 2 out of 5 people that purchased diaper also bought beer, the confidence is then 2/5 = 0.4.
 
 - Lift : The effect X has on the occurence of Y, comparing to the expect occurence of Y, for example, a lift of 4 indicates that given X, Y is 4 times more likely to appear than expected. 

In my opinion, lift is the most important metric, however, support and confidence also needs to be considered when generating rules. A rule with high lift but little support might be a coincidence or a flavor of some particular customer, and that a rule with high lift but little confidence means that the occurence of Y is still too unlikely comparing to other items, therefore hard to create value in business. With these in mind, my way of creating rules would be setting a minimum support and confidence, then sorting by lift to find the most important rules among all.  
 
We will choose a minimum support of 0.001 and a minimum confidence of 0.4, and setting the  maximum length of rules to be 5.
```{r,results='hide'}
rules <- apriori(trans, parameter = list(supp = 0.001, conf = 0.4,maxlen = 5))
rules<-sort(rules, by="lift", decreasing=TRUE)
options(digits = 3)
```

##Inspecting the rules

```{r,warning=FALSE}
inspect(rules[1:4])
summary(rules)
```
```{r}
plot(rules,method = 'scatterplot', measure = c("support","confidence"), shading = "lift")
```
From the summary and scatter plot of rules:
 
 - Most of the rules are length of 3.
 - Most of the rules have support between 0.001 and 0.002.
 - There seems to be a few rules that are very strong.

 
```{r}
plot(rules[1:20], method="graph",control = list(cex=0.5))
```
The graph above enable us to visualize the cluster of items. From the graph, we can see some interesting rules, just to mention a few, along with some new questions: 
 
 - Trivial : A person who buy Whole Milk Yogurt tend to purchase a bottle of Whole Milk, this makes sense to me. Does the other way around work the same and what can we do with this combination?
 
 - Mysterious : Fat Free Blueberry yogurt and Fat Free Strawberry yogurt, as well as 0% Blueberry Acai Greek Yogurt and 0% Raspberry Yogurt have an unusually strong lift and little support which seems strange to me. Does the relationship really exist or only because of some particular customers?
  
 - Useful : There is a group of items that is associated with limes, there seems to exist some unknown but useful relationship. Are they all correlated and does this phenomenon have any business potential?
 
##Answer the questions

#### 1. Confirm the trivial
```{r}
yogurt_milk<-apriori(data=trans, parameter=list(supp=0.001,conf = 0.4), 
               appearance = list(rhs="Organic Whole Milk",default="lhs"),
               control = list(verbose=F))
inspect(yogurt_milk)
support_list<-rules@quality$support
support_rank<-length(support_list[support_list>yogurt_milk@quality$support])
paste("The rule has the",support_rank,"largest support among 191 rules extracted")

lift_list<-rules@quality$lift
lift_rank<-length(lift_list[lift_list>yogurt_milk@quality$lift])
paste("The rule has the",lift_rank,"largest lift among 191 rules extracted")

confidence_list<-rules@quality$confidence
confidence_rank<-length(confidence_list[confidence_list>yogurt_milk@quality$confidence])
paste("The rule has the",confidence_rank,"largest confidence among 191 rules extracted")
```
This rule is a strong rule among all rules, with a nice support and a large lift over 10, however it has a slightly low confidence. Let's see if the other way around, {Whole Milk Plain Yogurt} => {Organic Whole Milk} works the same.

```{r}
milk_yogurt<-apriori(data=trans, parameter=list(supp=0.001,conf = 0.01), 
               appearance = list(rhs="Whole Milk Plain Yogurt",default="lhs"),
               control = list(verbose=F))
inspect(milk_yogurt)
```
Support and Lift are the same because they are symmetric. However, the confidence for the rule is only 0.04, indicating that buying Organic Whole Milk would not usually lead to buying Whole Milk Plain Yogurt. Therefore, a business strategy might be promoting Whole Milk Plain Yogurt with a little box of Oragnic Whole Milk as complementary.

#### 2. Test the mysterious
```{r}
fat_free_yogurt<-apriori(data=trans, parameter=list(supp=0.001,conf = 0.4), 
               appearance = list(rhs="Fat Free Strawberry Yogurt",default="lhs"),
               control = list(verbose=F))
inspect(fat_free_yogurt)
find_user_id<-function(id){
  cond<-all(c("Fat Free Strawberry Yogurt","Fat Free Blueberry Yogurt")%in%orders[orders$order_id==id,"product_name"])
  return(ifelse(cond,unique(orders[orders$order_id==id,"user_id"]),-1))
}
ids<-sapply(unique(orders$order_id),find_user_id)
ids<-ids[ids>=0]
paste("There are",length(unique(ids)),"unique users in",length(ids),"total users that purchased the combination")
```
It seems that there are all customers that bought the Fat Free combination are unique in the data time range. However, there might be other reasons that caused the high lift, such as an existing promotion or so, which needs further investigation. 

#### 3. Investigate the useful
```{r}
lime<-apriori(data=trans, parameter=list(supp=0.001,conf = 0.4), 
               appearance = list(rhs="Limes",default="lhs"),
               control = list(verbose=F))
inspect(lime)
```

From our rules, we can see a bunch of items that customers often purchase with Limes, they are Jalapeno Peppers, Cilantro, Avocado, Baby Spinach, Garlic, Asparagus, and Large Lemon. Feeding these names to google gave me some special dishes such as baby spinach avocado salad, Garlic Asparagus with Lime, or Cilantro Lime Asparagus and Rice as results. These are useful information that I don't know since I don't cook at all, and it seems that the rules do exist, and that Limes or Large Lemon are important ingredients for all these dishes. Therefore, a business strategy is to place these items close together, while putting Limes and Large Lemon far away from them, so that customers would have to walk through other department to get them, thus increasing the chance of other purchases.

##Conclusion
In this small project I only chose serveral rules to look into because of time, but there is a lot more rules to work with. This is a truely amazing dataset, thanks to Instacart, and there are many other things that can be done to the dataset, such as determining different purchase behavior of different time, or to build a simple recommendation system base on the rules and collaborative filtering between users. Maybe next time. 